Parallel multi-objective optimization using co-evolutionary islands and search space overlapping

### 2016-12-15 hpmoon-coor.tex
Added Antonio's paper and some SOA modifications. Closes #13

The hypothesis of this paper is that using overlapped sections of the chromosome, calculated depending the number of islands to use, in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time with respect to the baseline, disjoint and partly overlapped methods. To demonstrate this, a new overlapping islands scheme has been compared with a baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a coevolutionary disjoint section approach, and a fixed-size overlapped method. Different benchmarks problems and different large number of islands have been used. Results show that this new technique can improve different quality indicators in the same amount of time when the problem size is large.

The rest of the paper is organized as follows: after a background in parallelization in MOEAs, 
the compared algorithms and the used methodology are presented in Section \ref{sec:coevo}. %PABLO: no puedo usar \ref{sec:met} porque la plantilla de esta revista no lo coge :/ (y además no tienen número)
Then, the results of the experiments are shown (Section \ref{sec:res}), followed by conclusions and suggestions for future work lines.

@@

The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of solutions was divided into the islands using a transformation of coordinates. Their authors concluded that dividing the search space is a good idea, but achieve this is not trivial. Other approaches for dividing the search space include objective separation in different processors \citep{Xiao03specialized}, or divide in two populations: one for elite and other search \citep{Wang09parallel}. Martens, in \citep{Martens13asynchronous} generated a Barabasi-Albert network as the island topology, and selection based on migrating individuals from a not-crowded area and acceptance based on diversity.

More similar approaches to the one presented here were studied in the
next two works, also using cooperative coevolution for high-dimensional problems. 
Dorronsoro et al. \citep{Dorronsoro13superlinear} obtained super-linear performance
in several instances using cooperative coevolution, focusing each island on a portion of the chromosome. However, in some instances the parallel version 
improved the sequential version. Kimovski et al. \citep{Kimovski15Parallel} proposed a
master-slave method that splits the population into several processors,
each one running in parallel a MOEA that only affects some portion of
the individuals. After a certain number of generations, the master
process receives all the sub-populations to be combined. Different
combination alternatives in this step are compared. Up to 8 processors
were used in the experimental setup. Our approach in this paper does not
broadcast all solutions to all islands for recombination, as previous
works, but only one solution to a random island, needing less
communication time. Besides, the maximum number of islands of
Dorronsoro or Kimovsky approaches was 8, while in this paper we have
used up to 128 islands.

Lately, some researchers \citep{cheng2015adaptive} have focused on
working on solution space, dividing the Pareto front in different
clusters in order to model it. This {\em divide and conquer} approach
is readily paralellizable and, in the sense of giving the
responsibility of different parts of the search space to different
modules, is similar to the one presented in this paper.

% ¿Se podría mencionar por algún lado el trabajo en algoritmos de
% hormigas paralelos de Antonio? - JJ
In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two co-evolutionary dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome, while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, coevolution with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivates us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Moreover, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, we are validating the
previous methods.

@@

As in the baseline, every certain number of generations, an individual is migrated to another random island.  In the new island, this individual will be considered as one of the others in the island, crossed and mutated in the same way, depending on the island identifier. Note that, on the contrary of other works such as the ones described by Talbi et al. \citep{Talbi08Parallel} all the islands deal with complete chromosomes for fitness calculation, so our approach can deal with no decomposable problems. The proposed methods  are as follows:

@@

Pseudo-code of the proposed algorithms is described in Algorithm \ref{fig:pseudo}.

@@

\begin{table*}
\begin{center}
\begin{tabular}{|c|c|}
\hline
{\em Parameter Name} & {\em Value} \\ \hline
Number of islands ($P$) & 8, 16, 32, 64 and 128 \\ \hline
Chromosome size ($L$) & 512 and 2048 \\ \hline
Execution time (s) & 25 (for 512) and 100 (for 2048) \\ \hline \hline
Global population size ($N$) & 1024 \\ \hline
Selection type & Binary Tournament Selection \\ \hline
Replacement type & Generational \\ \hline 
Crossover type & SBX \\ \hline
Mutation  type & Polynomial\\ \hline
Mutation probability & 1/$L$ \\ \hline
Generations between migration & 5 \\ \hline
Individuals per migration & 1 \\ \hline
Selection for migration & Binary Tournament\\ \hline
Runs per configuration & 30 \\ \hline
\end{tabular}
\caption{Parameters and operators used in the experiments.}
\label{tab:parameters}
\end{center}
\end{table*}

>>>>

The hypothesis of this paper is that using overlapped sections of the chromosome, calculated depending the number of islands to use, in a coevolutionary multi-objective island-based algorithm can improve the quality of the solutions in the same computing time with respect to the baseline, disjoint and partly overlapped methods. To demonstrate this, different overlapping islands schemes have been compared: a baseline version of a distributed NSGA-II \citep{Deb00NSGAII}, a coevolutionary disjoint section approach version of it, a fixed-size overlapped version and our new adaptive method. Different benchmarks problems and different large number of islands have been used. Results show that this new technique can improve different quality indicators in the same amount of time when the problem size is large.

The rest of the paper is organized as follows: after the State of the Art in distributed and coevolutionary MOEAs, 
the used methodology and the compared algorithms are described in Section \ref{sec:coevo}. 
Then, the results of the experiments are presented (Section \ref{sec:res}), Finally, the conclusions and future work are discussed.

@@

The work by Deb et al. \citep{Deb03distributed} was one of the first approaches for distributed MOEAs (dMOEAs). In that work, the dominance of the solutions was divided into the islands using a coordinate transformation. Their authors concluded that dividing the search space is a good idea, although achieve this is not trivial. Dividing the search space has been explored by other researchers, for example, dividing the population in elite and search sub-populations \citep{Wang09parallel}, or separation in processors by objective \citep{Xiao03specialized}. Other authors, such as Martens \citep{Martens13asynchronous} use migration to accept individuals based on diversity, and migrating from not-crowded areas.

Addressing high-dimensional problems with cooperative coevolution has also been studied in several works with approaches more similar to the one presented here.
The approach to focus on a portion of the chromosome, as in our overlapped method, was firstly used in the work by Dorronsoro et al. \citep{Dorronsoro13superlinear}, obtaining super-linear performance
in several instances. This approach has also been used by Kimovski et al. \citep{Kimovski15Parallel}, but using a
master-slave method that splits the population into several processors. As in previous work, each node
runs a parallel MOEA that only affects some portion of 
the individuals and the master
process receives all the sub-populations to be combined every certain number of generations. Up to 8 processors
were used in the experimental setup, and several
combination alternatives were compared. The main difference of our work with respect to previous works is that our approach does not 
broadcast all solutions to all islands for recombination, but only one solution to a random island, needing less
communication time. Moreover, Dorronsoro or Kimovsky approaches limited the maximum number of islands to 8, while in this paper we have
used up to 128 islands.

Giving the responsibility of different parts of the search space to different modules, has also been explored lately by some researchers. For example, some authors focus on the solution space modeling the PF by dividing it in different clusters, allowing an easily parallelizable {\em divide and conquer} approach \citep{cheng2015adaptive}. Even other types of metaheuristics, such as the Multi-Objective Ant Colony Optimization (MOACO), can rely on island models to divide the search space depending on the node to achieve better results \cite{Mora13Moacos}.

% ¿Se podría mencionar por algún lado el trabajo en algoritmos de
% hormigas paralelos de Antonio? - JJ Fergu: DONE
In our previous work \citep{Garcia16hpmoon} we used some of
the aforementioned ideas to compare two co-evolutionary dMOEAs. The first
one divided the chromosome in $P$ sections, being $P$ the number of
islands. Every island $p$ only performed the mutation and crossover in
that part ($p_{th}$) of the chromosome, while the fitness was
calculated using the whole individual. After a certain number of
generations, individuals were migrated randomly to other islands. The
performance metrics were calculated at the end of the run. The second
method, coevolution with overlapping islands, used the $p_{th-1}$ and
$p_{th+1}$ sections of each chromosome, besides the $p_{th}$
part. During the same amount of time both methods obtained better
results than a baseline algorithm that dealt with the whole chromosome
in each island for crossover and mutation. We discovered that the
performance using one or another method depends on the number of
sections of the individuals and the number of islands used. This
motivates us to find a new adaptive method to select this number of
sections of the chromosome to use, depending on the number of
islands. Besides, previous experiments were performed on a single
processor synchronous island model with a limited number of islands
(8, 32 and 128). In this paper,  8, 16, 32, 64 and 128 islands have
been used, and this time, the experiments have been run on a real
cluster. Therefore, at the same time we are proposing an adaptive
method to divide the individual search space, we are validating the
previous methods.

@@

As in the baseline (B), an individual is migrated to another random island after a fixed number of generations. In the new island, this individual will be considered as one of the others in the island, crossed and mutated in the same way, depending on the island identifier. Note that, on the contrary of other works such as the ones described by Talbi et al. \citep{Talbi08Parallel} all the islands deal with complete chromosomes for fitness calculation, so our approach can deal with no decomposable problems. The proposed methods for our experimental section are as follows:

@@

Pseudo-code of the proposed algorithms is described in Algorithm \ref{alg:EA}.

@@

\begin{table*}
\begin{center}
\begin{tabular}{|c|c|}
\hline
{\em Parameter Name} & {\em Value} \\ \hline
Global population size ($N$) & 1024 \\ \hline
Selection type & Binary Tournament Selection \\ \hline
Replacement type & Generational \\ \hline 
Crossover type & SBX \\ \hline
Mutation  type & Polynomial\\ \hline
Mutation probability & 1/$L$ \\ \hline
Individuals per migration & 1 \\ \hline
Generations between migration & 5 \\ \hline
Selection for migration & Binary Tournament\\ \hline
Runs per configuration & 30 \\ \hline \hline
Number of islands ($P$) & 8, 16, 32, 64 and 128 \\ \hline
Chromosome size ($L$) & 512 and 2048 \\ \hline
Execution time (s) & 25 (for 512) and 100 (for 2048) \\ \hline \hline
\end{tabular}
\caption{Parameters and operators used in the experiments.}
\label{tab:parameters}
\end{center}
\end{table*}
